Relationships in the extremes and their influence on competition and coexistence

### 2021-07-07 paper/Paper.Rnw
small changes to compile

%Put other possible titles below, commented, and we will select the best one at the end
%Relationships between extreme events and their influence on competition and coexistence

\author[a]{Jasmin Albert}
\author[a,b]{Daniel C. Reuman}
%***DAN: Jasmin, you should choose one way your name will appear, exactly, on all your publications,
%so everyone knows its you. For instance, I always appear as Daniel C. Reuman, not Daniel Reuman, not
%Dan Reuman, etc. This makes it possible to find all my papers by searching. So please decide what
%name to use, Jasmin Albert or Pimsupa Albert or P. Jasmin Albert or whatever. Also keep in mind
%that because "Albert" is a common name you might want something less common for the rest of it 
%so that your papers are not confused with anyone elses, though of course I emphasize that how you
%are going to be known it totally up to you, I'm just recommending you think it through now so you
%can be consistent. I recognize that this all may be moot if you one day change your name and want to
%be known by the new name on papers after that point, but anyway in case you don't ever change your 
%name you might as well be consistent now, at least. So please change your name above to whatever
%you want to be officially known as on scientific papers. Same change in the sup mat, please.

@@

\begin{document}
%\SweaveOpts{concordance=TRUE}

%The following is where you load in the numeric results that will be embedded in the text
<<eval=T,echo=F,message=F,warning=F>>=
system(paste("biber", sub("\\.Rnw$", "", current_input())))

#place R code here for loading in necessary variables from Results
#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
allregres<-matrix(rnorm(10),2,5)
@

@@

Jasmin, I think the below makes a pretty good start at tryinbg to nail down a good way to express our goals
in the last para of the Intro, so please incorporate these into your last para. If you have a better way
to phrase these, please put both so we can compare at some point when we meet.
G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
models exhibiting solely temporal variation; spatial variation can be considered in future work.
G2) To explore using a simple model the relative magnitude of the contributions of coexistence that come
from ATA compared to other mechanisms.

@@

In this section we pursue goal G1 from the Introduction: to augment the MTC to show how ATA
plays a role in species coexistence.
%***DAN: make sure you define MTC and ATA once (probably in the Intro) and then use the abbreviations onle after that.
For pedgogical clarity, we develop each step of the theory in general and for a simple classical model 
(the lottery model), as an example.

The lottery model is as follows. 
%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year.
For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
properties, corresponding to various tail association cases. Note that $N$ is constant through time.

Modern coexistence theory \citep{Chesson_2000} and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
%***DAN: Search for REF and replace all with appropriate citations
quantify the contributions of 
different mechanisms to species coexistence. But we show below
that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
We now define storage effects for a general two species model, and specifically 
for the lottery model, using an approach and notation
similar to those of \cite{Ellner_2016} and \cite{Ellner_2019}; see those papers for additional details.

For a general model, it is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
of species $i$ can be written as an increasing function of
an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
competitive pressure on species $i$, $C_i(t)$. As for the lottery model example, we here assume, for simpicity, 
that the $E_i(t)$ are iid through time, and likewise for the
$C_i(t)$. Though this assumption can be circumvented, we make the assumption because it simplifies the
presentation of new ideas.

For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/\delta N$, which
does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
Species growth rates for the lottery model can then be written 
$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$.

In general, the storage effect for species $i$ is the contribution of covariation between $E$ and 
$C$ to a difference between invader and resident mean growth rates, when $i$ is the invader.
Specifically, storage effects for species $i$ are 
\begin{equation}
\Delta_i^{(EC)} = (\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i}),\label{eq:storage_effects_gen}
\end{equation}
where: $\olr_{i \bs i}$ refers to the mean ``invasion growth rate'' of species $i$, that is, the mean growth rate of 
species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;
$\olr_{j \bs i}$ refers to the mean steady-state 
growth rate of species $j$ when species $i$ is absent or at negligible abundance (this is $0$, \emph{a priori});
and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative modelling scenario, identical except that
influences of environment and competition on the invader or resident have been rendered independent.  
Following \cite{Ellner_2016,Ellner_2019}, a subscript 
$i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
The notation $\Delta_i^{(EC)}$ for storage effects also follows \cite{Ellner_2019}.
The \emph{scaling factor}
$q_{ij}$ measures the relative sensitivities to competition of the two species, see \cite{Chesson_1994},
\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the definition of $q_{ij}$. 
Since it was assumed that $r_i=r_i(E_i,C_i)$, we can write $\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})]$, where
$\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions. Also, $\olr_{j \bs i}=\E[r_j(E_j,C_{j \bs i})]$.
An expression for $\olrs_{i \bs i}$ is obtained from $\olr_{i \bs i}$ by replacing the bivariate random 
variable $(E_i,C_{i \bs i})$ by a random variable with the same marginal distributions, but independent components, 
$(E_i^\shp,C_{i \bs i}^\shp)$, so $\olrs_{i \bs i}=\E[r_i(E_i^\shp,C_{i \bs i}^\shp)]$.
Likewise, $\olrs_{j \bs i}=\E[r_j(E_j^\shp,C_{j \bs i}^\shp)]$.

For the lottery model, when $i$ is an invader, the competition experienced by either 
species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
Likewise, 
$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
though we already knew this equals $0$. 
An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
We show in SI section \ref{SI-SIsect:scaling_factors}
that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
%***DAN: Note this might be wrong. Some of the language in Ellner et al (2016) indicates q_{ij} here should be 1,
%even if \mu_1 and \mu_2 are distinct, as long as we use the same \delta for both species (which we do).
%But when I try to follow the definitions of the same paper to compute q_{ij} I get the quantity listed above,
%which is not 1. I'll need to get to the bottom of this, possibly by speaking to Ellner, and then put the right 
%value here, and also write the corresponding sup mat section accordingly (though you may not need that section, 
%it may be enough to just cite Ellner and Chesson if you go with q_ij=1.
Since we assumed log-normally distributed $B_i$, 
$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
%***DAN: This statement also needs adjusting if you go with q_ij=1
So $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into (\ref{eq:storage_effects_gen}).

We now specify how to decompose storage effects into
contributions due to asymmetric tail associations and due
to $EC$ correlation \emph{per se}. 
Storage effects for species $i$ were computed by comparing
differences between invader and resident mean growth rates for the original model and for a 
modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
Because these components were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole $EC$ relationship, in all its aspects, to the invader/resident growth rate
difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
and with some but not all aspects of the relationship 
between the components altered relative to the original distributions. Specifically, asymmetries of tail association are removed -- see below for how this is done. We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
\begin{align}
\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
&= \Delta_i^{[E \nat C]} + \Delta_i^{[EC]}.\label{eq:storage_effects_further}
\end{align}
The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to 
asymmetries of tail association between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the removal of part of the relationship between $E$ and $C$ components.
The term $\Delta_i^{[EC]}$ is the component due to $EC$ correlation \emph{per se}, after 
removing the effects of asymmetric tail associations. 
The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
this is a decompostion of storage effects, which were denoted by \cite{Ellner_2019} using a superscript $(EC)$,
i.e., $\Delta_i^{(EC)}$, to indicate that storage effects are, in turn, part of a decomposition of a difference between invader and
resident growth rates (see below).

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? We assume that the cummulative distribution functions (cdfs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the cdf of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal.  Let $\rho_i$ be the covariance of
these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\var(e_i)=\var(c_{i \bs i})=1$, and we denote $\cov(e_i,c_{i \bs i})$ by $\tilde{\rho}_i$. We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\tilde{\rho}_i$ is then determined 
in one of 
four alternative ways: $\tilde{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated in the sense
described in the Introduction.
%***DAN: This means you have to introduce this well, and explain why the way we define tail assoc is
%the *right* way
For those familar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\tilde{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created in an analogous way.

For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
$\tilde{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
can be written as $(\exp(\tilde{b}_i),\exp(\tilde{b}_j))$, where $(\tilde{b}_i,\tilde{b}_j)$
is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\tilde{b}_i)=\var(\tilde{b}_j)=\sigma$,
and $\cov(\tilde{b}_i,\tilde{b}_j)=\cov(b_i,b_j)$. So $(\tilde{b}_i,\tilde{b}_j)$ has the same
marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
$(b_i,b_j)$ itself need not be. Given these definitions, it is straightforward to compute the 
component $\olrn_{i \bs i}$ of (\ref{eq:storage_effects_further}) for the lottery model using the simulation 
approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
so $E_j^\nat=E_j$ and $C_{j \bs i}^\nat = C_{j \bs i}$. Thus $\olr_{j \bs i}=\olrn_{j \bs i}$.

>>>>

%Put other possible titles below, commented, and we will select the best one at the end
%Relationships between extreme events and their influence on competition and coexistence
%Coexistence and consequences of extreme environmental relationships
%Coexistence and consequences of environmental relationships in their extremes
%Coexistence and implications of asymmetric environmental relationships
%Extreme [climatic] events and their implications for biodiversity
%Deductions to coexistence, extreme events, and asymmetric relationships
%Extreme relationships (relationships in their extremes) and their deductions ([negative] contributions) to coexistence

\author[a]{Pimsupa Jasmin Albert}
\author[a,b]{Daniel C. Reuman}
%***DAN: Jasmin, you should choose one way your name will appear, exactly, on all your publications,
%so everyone knows its you. For instance, I always appear as Daniel C. Reuman, not Daniel Reuman, not
%Dan Reuman, etc. This makes it possible to find all my papers by searching. So please decide what
%name to use, Jasmin Albert or Pimsupa Albert or P. Jasmin Albert or whatever. Also keep in mind
%that because "Albert" is a common name you might want something less common for the rest of it 
%so that your papers are not confused with anyone elses, though of course I emphasize that how you
%are going to be known it totally up to you, I'm just recommending you think it through now so you
%can be consistent. I recognize that this all may be moot if you one day change your name and want to
%be known by the new name on papers after that point, but anyway in case you don't ever change your 
%name you might as well be consistent now, at least. So please change your name above to whatever
%you want to be officially known as on scientific papers. Same change in the sup mat, please.

@@

\begin{document}
\SweaveOpts{concordance=TRUE}
\SweaveOpts{concordance=TRUE}

%The following is where you load in the numeric results that will be embedded in the text
%<<eval=T,echo=F,message=F,warning=F>>=
%system(paste("biber", sub("\\.Rnw$", "", current_input())))

%#place R code here for loading in necessary variables from Results
%#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
%allregres<-matrix(rnorm(10),2,5)
@

@@

@@

For pedgogical clarity, we develop each step of the theory both for a simple classical model 
(the lottery model), and in general. The lottery model is as follows. 
%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year.
For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
properties, corresponding to various tail association cases. Note that $N$ is constant through time.

Modern coexistence theory \citep{Chesson_2000} and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
%***DAN: Search for REF and replace all with appropriate citations
quantify the contributions of 
different mechanisms to species coexistence. But we show below
that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
We now define storage effects for a general two species model, and specifically 
for the lottery model, using an approach and notation
similar to those of \cite{Ellner_2016} and \cite{Ellner_2019}; see those papers for additional details. 
%***DAN: Don't forget to return to the other mechanisms of coexistence below. Basically want to say intuitively 
%why tail associations don't influence those,
%though might be worth mentioning in the Discussion that events are getting more extreme. We have studied how *relationships*
%in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
%in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
%to also look at the other mechanisms. May still need a sup mat laying out all the math.

For a general model, it is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
of species $i$ can be written as an increasing function of
an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
competitive pressure on species $i$, $C_i(t)$. For simplicity, we here assume 
the $E_i(t)$ are iid through time, and likewise for the
$C_i(t)$. Though this assumption can be circumvented, we make the assumption because it simplifies the
presentation without obscuring ideas. 
For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/\delta N$, which
does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
Species growth rates for the lottery model can then be written 
$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$.

In general, the storage effect for species $i$ is the contribution of covariation between $E$ and 
$C$ to a difference between invader and resident mean growth rates, when $i$ is the invader.
Specifically, storage effects for species $i$ are 
\begin{equation}
\Delta I_i = (\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i}),\label{eq:storage_effects_gen}
\end{equation}
where: $\olr_{i \bs i}$ refers to the mean ``invasion growth rate'' of species $i$, that is, the mean growth rate of 
species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;
$\olr_{j \bs i}$ refers to the mean steady-state 
growth rate of species $j$ when species $i$ is absent or at negligible abundance (this is $0$, \emph{a priori});
and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative model scenario, identical except that
influences of environment and competition on the invader or resident have been rendered independent.  
A subscript 
$i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
The \emph{scaling factor}
$q_{ij}$ measures the relative sensitivities to competition of the two species, see \cite{Chesson_1994},
\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the definition of $q_{ij}$.

For the lottery model, when $i$ is an invader, the competition experienced by either 
species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
Here $\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions for which
an overbar would be cumbersome. Likewise, 
$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
though we already knew this equals $0$. An expression for $\olrs_{i \bs i}$ is then obtained from the 
above expression for $\olr_{i \bs i}$ by replacing the bivariate random variable $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise, a formula for 
$\olrs_{j \bs i}$ is available by replacing $(E_j,C_{j \bs i})$ in an analogous way. We use 
$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
We show in SI section \ref{SI-SIsect:scaling_factors}
that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
%***DAN: Note this might be wrong. Some of the language in Ellner et al (2016) indicates q_{ij} here should be 1,
%even if \mu_1 and \mu_2 are distinct, as long as we use the same \delta for both species (which we do).
%But when I try to follow the definitions of the same paper to compute q_{ij} I get the quantity listed above,
%which is not 1. I'll need to get to the bottom of this, possibly by speaking to Ellner, and then put the right 
%value here, and also write the corresponding sup mat section accordingly (though you may not need that section, 
%it may be enough to just cite Ellner and Chesson if you go with q_ij=1.
Since we use log-normally distributed $B_i$, 
$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
%***DAN: This statement also needs adjusting if you go with q_ij=1
So $\Delta I_i$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into (\ref{eq:storage_effects_gen}).

Having just reviewed storage effects, we now specify how to decompose storage effects into
contributions due to asymmetric tail associations and due
to $EC$ correlation \emph{per se}. 
Storage effects for species $i$ were computed by comparing
differences between invader and resident mean growth rates for the original model and for a 
modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
Because these components were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole $EC$ relationship, in all its aspects, to the invader/resident growth rate
difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
and with some but not all aspects of the relationship 
between the components altered relative to the original distributions. Specifically, asymmetries of tail association are removed -- see below for how this is done. We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
\begin{align}
\Delta I_i &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \\
&= I_i^{(\text{ata})} + I_i^{(\text{cps})}.\label{eq:storage_effects_further}
\end{align}
The superscript ``ata" stands for ``asymmetric tail associations'' because this component of storage effects is due to 
asymmetries of tail association between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$. 
The superscript ``cps" stands for ``correlation
\emph{per se}'' because this component is due to $EC$ correlation \emph{per se}, after 
removing the effects of asymmetric tail associations.
The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the removal of part of the relationship between $E$ and $C$ components.

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? We assume that the cummulative distribution functions (cdfs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the cdf of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal.  Let $\rho_i$ be the covariance of
these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\var(e_i)=\var(c_{i \bs i})=1$, and denote $\cov(e_i,c_{i \bs i})$ by $\tilde{\rho}_i$. We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\tilde{\rho}_i$ is then determined 
in one of 
four alternative ways: $\tilde{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated in the sense
described in the Introduction.
%***DAN: This means you have to introduce this well, and explain why the way we define tail assoc is
%the *right* way
For those familar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\tilde{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created in an analogous way.

For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
$\tilde{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
can be written as $(\exp(\tilde{b}_i),\exp(\tilde{b}_j))$, where $(\tilde{b}_i,\tilde{b}_j)$
is bivariate normally distributed with mean $(\mu_i,\mu_j)$, $\var(\tilde{b}_i)=\var(\tilde{b}_j)=\sigma$,
and $\cov(\tilde{b}_i,\tilde{b}_j)=\cov(b_i,b_j)$. So $(\tilde{b}_i,\tilde{b}_j)$ has the same
marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
$(b_i,b_j)$ itself need not be. Given these definitions, it is straightforward to compute the 
components of (\ref{eq:storage_effects_further}) for the lottery model using the simulation 
approach of \cite{Ellner_2016,Ellner_2019}, though in the next section we describe some tricks
that make these computations efficient.

The other mechanism in modern coexistence theory that apply in our modelling context (i.e., purely temporal variation)
make no reference to covariation between quantities, whereas asymmetries of tail associations are 
defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
for which asymmetries of tail association are relevant. Details are given in SI section 
\ref{SIsect:other_mechanisms}, and we return briefly to related ideas in the Discussion.
%***DAN: In the Discussion, I intend to talk about how possible changes in tail association 
%patterns of env vars in nature may influence coexistence of species. It won't influence the 
%other mechanisms, though changes in the extreme values of environmental variables, known
%to be occurring, might.
